Real-space computation of $E$/$B$-mode maps I: Formalism and Compact Kernels

### 2018-03-30 articles/formalism/sections/generalizedeb.tex
updates to the polarization primer section of the formalism paper

We can characterize different forms of the radial kernel by introducing the following harmonic space operator,
%
\beq
\tilde{\mathcal{G}} = {\begin{bmatrix} g_{\ell}^E & 0  \\  0 & g_{\ell}^B \end{bmatrix}} \,,
\eeq
%
where the functions $g_{\ell}^E$ and $g_{\ell}^B$ represent the harmonic representation of the modified radial functions and can in the most general case be chosen to be different for $E$ and $B$ modes. To simplify the discussion and without loosing generality we proceed by setting $g_{\ell}^E = g_{\ell}^B= g_{\ell}$. Once we have made a choice for these harmonic functions, we can define the real space operator $\bar{O}'$ which translates Stokes $Q$ \& $U$ to scalars $E$ \& $B$ and the inverse operator $\bar{O}'^{-1}$ in the following manner,
%
\begin{subequations} \label{eq:gen_qu2eb}
\beqry
{\bar O}' &=& {{}_0\mathcal{Y}} *\tilde T^{-1}*\tilde{\mathcal{G}}* {{}_2\mathcal{Y}^{\dagger}} *\bar T \,,\\
{\bar O}'^{-1}&=& \bar{T}^{-1} *{{}_2\mathcal{Y}}* \tilde{\mathcal{G}}^{-1} *\tilde T *{{}_0\mathcal{Y}^{\dagger}}
\eeqry
\end{subequations}
%
where we have used the primed notation  to distinguish these generalized operators from the default operators defined in \sec{sec:qu2eb} and \sec{sec:eb2qu}. Note that for an arbitrary choice of $\tilde{\mathcal{G}}$ only one of the operators in \eq{eq:gen_qu2eb} is well defined, since $\tilde{\mathcal{G}}^{-1}$ may be ill defined. If we require both the forward and inverse operators to be well defined, then we are constrained in choosing $\tilde{\mathcal{G}}$ such that it has a valid  inverse. The radial part of these generalized convolution kernels is given by the following expressions,
%
\begin{subequations}
\beqry
G_{QU \rightarrow EB}(\beta) &=& G(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \, \label{eq:mod_rad_forward} \\
G_{EB \rightarrow QU}(\beta) &=& G^{-1}(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}^{-1}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \,,\label{eq:mod_rad_inverse}
\eeqry
\end{subequations}
%
where $g_{\ell}$ are the same multipole function as those appearing in $\tilde{\mathcal{G}}$. Given this general definition for the radial function $G(\beta)$, note that the default radial function ${{}_{\mm}f}$ is just a special case resulting from the choice $\tilde{\mathcal{G}}=\mathbb{1}$ ($g_{\ell}=1$). Note that for this choice of $\tilde{\mathcal{G}}$ the inverse is trivial $\tilde{\mathcal{G}}^{-1}=\tilde{\mathcal{G}}$ and therefore $G^{-1}(\beta) = G(\beta)$.

While defining these generalized operators, it seems more natural to choose the real space function $G(\beta)$ as compared to choosing the multipole function $g_{\ell}$. Using the orthogonality property of associated Legendre polynomials it can be shown that the harmonic function $g_{\ell}$ is given by the following integral over the radial function $G(\beta)$,
%
\beq
g_{\ell} = 2 \pi \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} \int _{0}^{\pi} G(\beta) P_{\ell}^{2}(\cos{\beta}) d\cos{\beta} \,. \label{eq:gb2bl}
\eeq
%
Here it is important to note that the radial function $G(\beta)$ has to be chosen such that it vanishes at $\beta=0$ and $\beta=\pi$. 
%One way to understand this is that the associated Legendre polynomials $P_{\ell}^2 \propto \sin^2{\beta}$ vanish at these values of the abscissa and hence cannot be used to describe functions which don't have this property.  Another way to understand this requirements is that at these locations the coordinate dependence of the Stokes parameters cannot be integrated out, since the azimuthal angle is ill defined and hence the convolution kernel needs to have vanishing contribution from these locations. \revisit{The normalization of these functions is not critically important, since any choice defines a convention. It is important to ensure consistency with the convention once a choice has been made.} \comment{find a better location for this}
Note that in contrast to the radial function $G(\beta)$ an instrumental beam function appropriately normalized has the property $B(\beta) \rightarrow 1$ as $\beta \rightarrow 0$. We clarify that the $B(\beta)$ refers to the effective beam acting to smooth the scalar $E$ \& $B$ mode maps. A circularly symmetric beam $B(\beta)$ defined at the pole can be expressed in the Legendre polynomial $P_{\ell}^0$ basis as follows,
%
\beq
B(\beta) = \sum_{\ell=0}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} b_{\ell} P_{\ell}^{0} (\cos{\beta})\,,
\eeq
%
where $b_{\ell}$ denote the coefficients of expansion.
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.48\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}\\
\subfigure[\label{fig:bl_gbeta}]{\includegraphics[width=0.48\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.48\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Left:} The vertical dashed gray line depicts the approximate pixel size $\Delta_{\rm pix} = \sqrt{\frac{4 \pi}{N_{\rm pix}}}$ of a Nside=512 Healpix map. The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arcminutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{N} \beta^n \exp{\left[ -\left( \frac{\beta-\beta0}{\sqrt{2} \sigma} \right)^s \right]}$ with parameters set to the following values $[n=1;\, \beta_0=0 ;\, \sigma = 2\Delta_{\rm pix} ;\, s=1.5]$. The black dashed curve depicts the band limited reconstruction of the modified radial function $G(\beta)$. We intentionally have plotted $\beta G(\beta)$ to clearly depict the high $\beta$ behavior of these functions. \textit{Middle: } This figure depicts the harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Right:} This figure depicts the beam function $B(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an instrument beam.}
\label{fig:example_gbeta}
\end{figure}
%

@@

\subsection{Recovering the default $E$ and $B$ mode spectra}
The generalized convolution kernels defined in the previous section, when operated on the Stokes vector returns some scalar $E'$ and $B'$ mode maps,
%
\beq
\bar{S}' = \bar{O}' * \bar{P}
\eeq
%
which are not the standard $E$ and $B$ modes maps. Since the the harmonic representation $g_{\ell}$ of the radial function $G(\beta)$ can be simply interpreted as the harmonic coefficients of some beam,  the spectra of the scalar fields $E'$ and $B'$ are related to the spectra of the standard $E$ and $B$ fields via the following relation, 
 %
 \begin{subequations}
 \beqry
C_{\ell}^{EE,BB,EB} &= &C_{\ell}^{E'E',B'B',E'B'} /   g_{\ell}^2\,,\\
C_{\ell}^{TE,TB}  &=&  C_{\ell}^{TE',TB'} / g_{\ell}\,,
 \eeqry
 \end{subequations}
 %
 where $C_{\ell}$ denotes the angular power spectra and $T$ refers to the temperature anisotropy map. Therefore the standard $E$ and $B$ mode spectra can be recovered from the modified fields $E'$ and $B'$ and their accurate recovery only relies on the inverse of the harmonic functions $1/g_{\ell}$ being well behaved, which can be ensured by making a suitable choice for the radial function $G(\beta)$.
%One direct application of this freedom in choosing the radial kernel is that one can mitigate the issue of leakage of power from $E$ to $B$.  By constructing a radial function which tapers to zero at sufficiently small radii one can clearly discard pixels which are affected by masking. 
%--------------------------------------------------------
%--------------------------------------------------------

>>>>

We can characterize different forms of the radial kernel by introducing the following harmonic space operator,
%
\beq
\tilde{\mathcal{G}} = {\begin{bmatrix} g_{\ell}^E & 0  \\  0 & g_{\ell}^B \end{bmatrix}} \,,
\eeq
%
where the functions $g_{\ell}^E$ and $g_{\ell}^B$ represent the harmonic representation of the modified radial functions and can in the most general case be chosen to be different for $E$ and $B$ modes. To simplify the discussion and without loosing generality we proceed by setting $g_{\ell}^E = g_{\ell}^B= g_{\ell}$. Once we have made a choice for these harmonic functions, we can define the real space operator $\bar{O}'$ which translates Stokes $Q$ \& $U$ to scalars $E$ \& $B$ and the inverse operator $\bar{O}'^{-1}$ in the following manner,
%
\begin{subequations} \label{eq:gen_qu2eb}
\beqry
{\bar O}' &=& {{}_0\mathcal{Y}} *\tilde T^{-1}*\tilde{\mathcal{G}}* {{}_2\mathcal{Y}^{\dagger}} *\bar T \,,\\
{\bar O}'^{-1}&=& \bar{T}^{-1} *{{}_2\mathcal{Y}}* \tilde{\mathcal{G}}^{-1} *\tilde T *{{}_0\mathcal{Y}^{\dagger}}
\eeqry
\end{subequations}
%
where we have used the primed notation  to distinguish these generalized operators from the default operators defined in \sec{sec:qu2eb} and \sec{sec:eb2qu}. Note that for an arbitrary choice of $\tilde{\mathcal{G}}$ only one of the operators in \eq{eq:gen_qu2eb} is well defined, since $\tilde{\mathcal{G}}^{-1}$ may be ill defined. If we require both the forward and inverse operators to be well defined, then we are constrained in choosing $\tilde{\mathcal{G}}$ such that it has a valid  inverse. As we will see this will be an important criteria to recover the standard CMB spectra. The radial part of these generalized convolution kernels is given by the following expressions,
%
\begin{subequations}
\beqry
G_{QU \rightarrow EB}(\beta) &=& G(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \, \label{eq:mod_rad_forward} \\
G_{EB \rightarrow QU}(\beta) &=& G^{-1}(\beta) = \sum _{\ell=2} ^{\ell_{\rm max}} g_{\ell}^{-1}\frac{2 \ell+1}{4 \pi} \sqrt{\frac{(\ell-2)!}{(\ell + 2)!}} P_{\ell}^2(\cos{\beta}) \,,\label{eq:mod_rad_inverse}
\eeqry
\end{subequations}
%
where $g_{\ell}$ are the same multipole function as those appearing in $\tilde{\mathcal{G}}$. Given this general definition for the radial function $G(\beta)$, note that the default radial function ${{}_{\mm}f}$ is just a special case resulting from the choice $\tilde{\mathcal{G}}=\mathbb{1}$ ($g_{\ell}=1$). Note that for this choice of $\tilde{\mathcal{G}}$ the inverse is trivial $\tilde{\mathcal{G}}^{-1}=\tilde{\mathcal{G}}$ and therefore $G^{-1}(\beta) = G(\beta)$.

While defining these generalized operators, it seems more natural to choose the real space function $G(\beta)$ as compared to choosing the multipole function $g_{\ell}$. Using the orthogonality property of associated Legendre polynomials it can be shown that the harmonic function $g_{\ell}$ is given by the following integral over the radial function $G(\beta)$,
%
\beq
g_{\ell} = 2 \pi \sqrt{\frac{(\ell-2)!}{(\ell+2)!}} \int _{0}^{\pi} G(\beta) P_{\ell}^{2}(\cos{\beta}) d\cos{\beta} \,. \label{eq:gb2bl}
\eeq
%
To ensure that the resultant field is a spin-0 field, it is important to note that the radial function $G(\beta)$ has to be necessarily chosen such that it vanishes at $\beta=0$ and $\beta=\pi$. 
%One way to understand this is that the associated Legendre polynomials $P_{\ell}^2 \propto \sin^2{\beta}$ vanish at these values of the abscissa and hence cannot be used to describe functions which don't have this property.  Another way to understand this requirements is that at these locations the coordinate dependence of the Stokes parameters cannot be integrated out, since the azimuthal angle is ill defined and hence the convolution kernel needs to have vanishing contribution from these locations. \revisit{The normalization of these functions is not critically important, since any choice defines a convention. It is important to ensure consistency with the convention once a choice has been made.} \comment{find a better location for this}

An arbirtrary $G(\beta)$ for which $g_{\ell} \neq 1$ can be equivalently thought in terms of the standard E and B mode fields being convolved with some circularly symmetric instrument beam. 
Note that in contrast to the radial function $G(\beta)$ an instrumental beam function appropriately normalized has the property $B(\beta) \rightarrow 1$ as $\beta \rightarrow 0$. We clarify that the $B(\beta)$ refers to the effective beam acting to smooth the scalar $E$ \& $B$ mode maps. A circularly symmetric beam $B(\beta)$ defined at the pole can be expressed in the Legendre polynomial $P_{\ell}^0$ basis as follows,
%
\beq
B(\beta) = \sum_{\ell=0}^{\ell_{\rm max}} \frac{2 \ell+1}{4 \pi} b_{\ell} P_{\ell}^{0} (\cos{\beta})\,,
\eeq
%
where $b_{\ell}$ denote the coefficients of expansion. \comment{Note that one can directly get E and B mode maps resulting from convolution with a non-circular beam by chosing the function $g_{\ell}$ to have some m dependence.}
%
\begin{figure}[!t] 
\centering
\subfigure[\label{fig:bl_gbeta}]{\includegraphics[width=0.48\columnwidth]{Gbeta_for_different_gl_lmax1536.pdf}}
\subfigure[\label{fig:glbl}]{\includegraphics[width=0.48\columnwidth]{gl_for_different_gbeta_lmax1536.pdf}}
\subfigure[\label{fig:gl_bbeta}]{\includegraphics[width=0.48\columnwidth]{Bbeta_for_gl_lmax1536.pdf}}
\caption{\textit{Left:} The vertical dashed gray line depicts the approximate pixel size $\Delta_{\rm pix} = \sqrt{\frac{4 \pi}{N_{\rm pix}}}$ of a Nside=512 Healpix map. The green line depicts the default radial kernel $f(\beta)$ defined in \eq{eq:qu2eb_gen_kernel}. The blue and orange lines depict the modified radial function resulting the beam harmonics $b_{\ell}$ corresponding to Gaussian beams with fwhm=15 \& 12 arcminutes respectively. The red curve depicts an example modified radial function: $G(\beta)=\mathcal{N} \beta^n \exp{\left[ -\left( \frac{\beta-\beta0}{\sqrt{2} \sigma} \right)^s \right]}$ with parameters set to the following values $[n=1;\, \beta_0=0 ;\, \sigma = 2\Delta_{\rm pix} ;\, s=1.5]$. The black dashed curve depicts the band limited reconstruction of the modified radial function $G(\beta)$. We intentionally have plotted $\beta G(\beta)$ to clearly depict the high $\beta$ behavior of these functions. \textit{Middle: } This figure depicts the harmonic representation of the respective radial functions as indicated by the legend. The dashed curves of the corresponding color depict the inverse of the harmonic functions. \textit{Right:} This figure depicts the beam function $B(\beta)$ evaluated from interpreting the respective harmonic functions as those corresponding to an instrument beam.}
\label{fig:example_gbeta}
\end{figure}
%

@@

\subsection{Recovering the default $E$ and $B$ mode spectra}
The generalized convolution kernels defined in the previous section, when operated on the Stokes vector returns some scalar $E'$ and $B'$ mode maps,
%
\beq
\bar{S}' = \bar{O}' * \bar{P}
\eeq
%
which as we show are mere filtered version of he standard $E$ and $B$ modes maps. The harmonic representation $g_{\ell}$ of the radial function $G(\beta)$ can be simply interpreted as the harmonic coefficients of some azimuthally symmetric beam $B(\beta) = \sum_{\ell} \frac{2 \ell+1}{4 \pi} g_{\ell} P^0_{\ell}(\cos{\beta})$. The spectra of the scalar fields $E'$ and $B'$ derived using the real space operators constructed using an arbitrary radial function $G(\beta)$ are related to the spectra of the standard $E$ and $B$ fields via the following relation, 
 %
 \begin{subequations}
 \beqry
C_{\ell}^{EE,BB,EB} &= &C_{\ell}^{E'E',B'B',E'B'} /   g_{\ell}^2\,,\\
C_{\ell}^{TE,TB}  &=&  C_{\ell}^{TE',TB'} / g_{\ell}\,,
 \eeqry
 \end{subequations}
 %
 where $C_{\ell}$ denotes the angular power spectra and $T$ refers to the temperature anisotropy map. Therefore the standard $E$ and $B$ mode spectra can be recovered from the modified fields $E'$ and $B'$ and their accurate recovery only relies on the inverse of the harmonic functions $1/g_{\ell}$ being well behaved, which can be ensured by making a suitable choice for the radial function $G(\beta)$. Examples of various forms of $G(\beta)$, its harmonic representation $g_{\ell}$ (and its inverse $1/g_{\ell}$ ) and the corresponding beam $B(\beta)$ are depicted in \fig{fig:example_gbeta}.
%One direct application of this freedom in choosing the radial kernel is that one can mitigate the issue of leakage of power from $E$ to $B$.  By constructing a radial function which tapers to zero at sufficiently small radii one can clearly discard pixels which are affected by masking. 
%--------------------------------------------------------
%--------------------------------------------------------

### 2018-03-30 articles/formalism/sections/real_space_formalism.tex
updates to the polarization primer section of the formalism paper

%--------------------------------------------------------
%--------------------------------------------------------
%\section{The spin-0 description of CMB polarization} \label{sec:pol-intro}
%--------------------------------------------------------
\section{Polarization primer}\label{sec:pol-primer}
The CMB polarization is measured in terms of Stokes Q and U parameters. These measurements can be combined to form the complex spin 2 polarization field as follows,
%
\beqry \label{eq:spin-pol}
_{\pm 2}\bar{X}(\hat{n}) &=& Q(\hat{n}) \pm i U (\hat{n}) \nonumber \\ &=& \sum_{\ell m}  {_{\pm 2}} \tilde{X}_{\ell m}  \, {_{\pm 2}}Y_{\ell m} (\hat{n}) \,.
\eeqry
%
Since these measured quantities depend on the local coordinate system, it is cumbersome to work with them. To overcome this, one describes the CMB polarization field in terms of a scalar field denoted by $E(\hat{n}) $ and a pseudo scalar field $B(\hat{n}) $ \cite{Kamionkowski1997}. These scalar fields are related to the spin-2 polarization field $_{\pm 2}X(\hat{n})$ via the following relations,
%
\beq \label{eq:ebdef}
\mathcal{E}(\hat{n}) = -\frac{1}{2} \big[ \bar{\eth}^2 _{+ 2}\bar{X}(\hat{n})  +  \eth^2 _{- 2}\bar{X}(\hat{n}) \big] ~\,;~\mathcal{B}(\hat{n}) = -\frac{1}{2i} \big[ \bar{\eth}^2 _{+ 2}X(\hat{n})  -  \eth^2 _{- 2}X(\hat{n}) \big] \,,
\eeq
%
where $\eth$ and $\bar{\eth}$ denote the spin raising and lowering operators respectively. These $E$ and $B$ fields are spin-0 fields similar to the temperature anisotropies and hence their value are independent of the coordinate system definitions (except that the B-modes have an odd parity, meaning that they change sign under reflection $\hat{n} \rightarrow -\hat{n}$.). The spin raising and lowering operators have the following properties \cite{goldberg67},
%
\beqrys \label{eq:spinopylm} 
\eth _s Y_{lm}(\hat{n}) &=& \sqrt{(\ell-s)(\ell+s+1)} _{s+1} Y_{lm}(\hat{n}) \,, \\
\bar{\eth} _s Y_{lm}(\hat{n}) &=& -\sqrt{(\ell+s)(\ell-s+1)} _{s-1} Y_{lm}(\hat{n}) \,, 
\eeqrys
%
where $_s Y_{lm}(\hat{n}) $ denote the spin-s spherical harmonics.

@@

For this work, we believe it is convenient to work in the z-y1-z2 convention, therefore all future references to Euler angles are to be assumed to be given in this convention. \footnote{The Euler angles in the more standard $z-y-z$ convention are related to those in the $z-y1-z2$ convection by the following rule: $(\alpha,\beta,\gamma)_{z-y-z} =(\gamma,\beta,\alpha)_{z-y1-z2}$ \cite{varshalovich}.}
In the z-y1-z2 convention, $\alpha$ defines the rotation about the z-axis, $\beta$ defines the rotation about the new y-axis (y1-axis) after the previous rotation and $\gamma$ defines the rotation about the final z-axis (z2-axis) after carrying out the previous two rotations. In more \revisit{physical terms}, these angles can be understood as follows: The rotation by $\alpha$ about the z-axis is such that it aligns the x-axis of the cartesian system at location $\hat{n}_0$ along the great circle in the direction of $\hat{n}_i$.  The rotation by $\beta$ about the y2-axis parallel transports the local cartesian coordinate system from location $\hat{n}_0$ to location $\hat{n}_i$, such that the z-axes of the two coordinate systems are parallel to each other. Finally the rotation by angle $\gamma$ about the z2-axis aligns the x \& y axes of the parallel transported system with those of the local cartesian system defined at $\hat{n}_i$. The different Euler angles and the corresponding rotations are depicted in \fig{fig:euler_angles}.

The simplest scenario is when one of the coordinates coincides with the north pole $\hat{n}_0=(\theta_0,\phi_0)=(0,0)$ (the local cartesian system at this point being defined with an arbitrary convention set by choosing $\theta_0 \rightarrow 0$ while moving along the longitude $\phi_0=0$), in which case it is easy to see that rotations by Euler angles: $(\alpha,\beta,\gamma) =(\phi_i,\theta_i,0)$ will align the cartesian system at $\hat{n}_0$ with that defined at position $\hat{n}_i$. 

@@

 where the terms on either side of the equation are those that appear in \eq{eq:qu2eb_ker_1}. Note that each sub-matrix in \eq{eq:qu2eb_ker_1} is formed by some linear combination of a complex number and its conjugate. The $m$ sum over the product of two spherical harmonic functions with spins $s_1$ and $s_2$ respectively, is given by the following identity \cite{varshalovich},

@@

Specifically note that $\alpha_{qe}, ~\beta_{qe} ~\&~ \gamma_{qe}$ denote the Euler angles which rotate the local cartesian system at the location $\hat{n}_q$ where the Stokes field is measured to the cartesian system at the location $\hat{n}_e$ where the scalar fields are evaluated: $\hat{n}_q \xrightarrow{\mathcal{R}(\alpha_{qe},\beta_{qe},\gamma_{qe})} \hat{n}_e$. The Euler angles for the inverse rotations are given by the Euler angles $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$\cite{varshalovich}. Since the kernels only depend on the cosine of the Euler angle $\beta$, they are immune to the change in its sign. More importantly, the ordering of indices  on the Euler angle $\alpha$ and $\gamma$  is important and leads to two different interpretations of the real space kernels.

@@

\bmat E_e \\ B_e  \emat_{q} =- {{}_{\mm}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{q e}) & \sin(2\alpha_{q e})\\  -\sin(2 \alpha_{q e})  & \cos(2 \alpha_{q e}) \emat  \bmat Q_{q} \\ U_{q}  \emat \Delta \Omega \,.

>>>>

%--------------------------------------------------------
%--------------------------------------------------------
%\section{The spin-0 description of CMB polarization} \label{sec:pol-intro}
%--------------------------------------------------------
\section{Polarization primer}\label{sec:pol-primer}
The CMB polarization is measured in terms of Stokes Q and U parameters. The Stokes Q is defined as the linear polarization measured along local cartesian axes ($+$ along the x-axis and $-$ along the y-axis) while Stokes U is defined as the linear polarization measured along the cartesian axes rotated by $45^{\circ}$ ($+$ along x-axis and $-$ along the y-axis in the counter-clockwise rotated system). Clearly the definitions of Stokes parameters depend on the choice of the coordinate system and it can be shown that the measurement made in two different coordinate systems rotated by an angle $\phi$ are related by the following equation,
%
\beq \label{eq:qu-rot}
\fqu' = \begin{bmatrix} \cos{2 \psi} &  \sin{2 \psi} \\ -\sin{2\psi} & \cos{2 \psi} \end{bmatrix} \fqu \,,
\eeq
%
which one can simply checked by verifying that $Q' \rightarrow U$ and $U' \rightarrow -Q$ when the coordinate systems are rotated by $45^\circ$ in the counter-clockwise sense.

Following the notation of \cite{Zaldarriaga1997}, fields which transform as,
%
\beq
{}_{s}f' = e^{-is\psi} {}_{s}f \,,
\eeq
%
under right handed rotations of the local coordinate system are termed spin-s fields. The Stokes parameters can be combined to define the complex fields,
%
\beq \label{eq:spin-pol}
_{\pm 2}\bar{X}(\hat{n}) = Q(\hat{n}) \pm i U (\hat{n}) \,, %\nonumber \\ &=& \sum_{\ell m}  {_{\pm 2}} \tilde{X}_{\ell m}  \, {_{\pm 2}}Y_{\ell m} (\hat{n}) \,.
\eeq
%
which under rotation of the local tangent plane coordinate system by an angle $\psi$ at any point on the sphere transform as: $_{\pm 2}X' = e^{\mp i2\psi} {_{\pm 2}X}$, which follows directly from \eq{eq:qu-rot}. Hence the quantities $_{\pm2}X$ behave as spin${\pm2}$ respectively.

The Stokes parameters being coordinate dependent quantities, it is cumbersome to work with them. It is possible to define spin raising ($\eth$) and spin lowering $(\bar{\eth})$ operators on the sphere \cite{goldberg67}.  These operators are such that when the spin raising operator acts on a spin-s field $_{s}f$ results in a fields with spin-$(s+1)$: $(\eth _{s}f)' = e^{-i(s+1)\psi}(\eth _{s}f)$ and similarly when a spin lowering operator acts on the field results in a field with spin-$(s-1)$: $(\bar{\eth} _{s}f)' = e^{-i(s-1)\psi}(\bar{\eth} _{s}f)$. This challenge is overcome by describing the CMB polarization field in terms of coordinate independent scalar fields denoted by $E(\hat{n}) $ and a pseudo scalar field $B(\hat{n}) $ \cite{Kamionkowski1997}. These scalar fields are constructed from the spin-2 polarization fields $_{\pm 2}X(\hat{n})$ via the following mathematical operations,
%
\beq \label{eq:ebdef}
\mathcal{E}(\hat{n}) = -\frac{1}{2} \big[ \bar{\eth}^2 _{+ 2}\bar{X}(\hat{n})  +  \eth^2 _{- 2}\bar{X}(\hat{n}) \big] ~\,;~\mathcal{B}(\hat{n}) = -\frac{1}{2i} \big[ \bar{\eth}^2 _{+ 2}X(\hat{n})  -  \eth^2 _{- 2}X(\hat{n}) \big] \,,
\eeq
%
where $\eth$ and $\bar{\eth}$ denote the spin raising and lowering operators respectively. These $\mathcal{E}$ and $\mathcal{B}$ fields are spin-0 fields similar to the temperature anisotropies and hence their value are independent of the coordinate system definitions (except that the B-modes have an odd parity, meaning that they change sign under reflection $\hat{n} \rightarrow -\hat{n}$.). The spin raising and lowering operators have the following properties \cite{goldberg67},
%
\beqrys \label{eq:spinopylm} 
\eth _s Y_{lm}(\hat{n}) &=& \sqrt{(\ell-s)(\ell+s+1)} _{s+1} Y_{lm}(\hat{n}) \,, \\
\bar{\eth} _s Y_{lm}(\hat{n}) &=& -\sqrt{(\ell+s)(\ell-s+1)} _{s-1} Y_{lm}(\hat{n}) \,, 
\eeqrys
%
where $_s Y_{lm}(\hat{n}) $ denote the spin-s spherical harmonics.

@@

For this work, it is convenient to think in the z-y1-z2 convention therefore all future references to Euler angles are in this convention. \footnote{The Euler angles in the more standard $z-y-z$ convention are related to those in the $z-y1-z2$ convection by the following rule: $(\alpha,\beta,\gamma)_{z-y-z} =(\gamma,\beta,\alpha)_{z-y1-z2}$ \cite{varshalovich}.}
In the z-y1-z2 convention, $\alpha$ defines the rotation about the z-axis, $\beta$ defines the rotation about the new y-axis (y1-axis) after the previous rotation and $\gamma$ defines the rotation about the final z-axis (z2-axis) after carrying out the previous two rotations. In more \revisit{physical terms}, these angles can be understood as follows: The rotation by $\alpha$ about the z-axis is such that it aligns the x-axis of the cartesian system at location $\hat{n}_0$ along the great circle in the direction of $\hat{n}_i$.  The rotation by $\beta$ about the y2-axis parallel transports the local cartesian coordinate system from location $\hat{n}_0$ to location $\hat{n}_i$, such that the z-axes of the two coordinate systems are parallel to each other. Finally the rotation by angle $\gamma$ about the z2-axis aligns the x \& y axes of the parallel transported system with those of the local cartesian system defined at $\hat{n}_i$. The different Euler angles and the corresponding rotations are depicted in \fig{fig:euler_angles}.

The simplest scenario is when one of the coordinates coincides with the north pole $\hat{n}_0=(\theta_0,\phi_0)=(0,0)$ (the local cartesian system at this point being defined with an arbitrary convention set by choosing $\theta_0 \rightarrow 0$ while moving along the longitude $\phi_0=0$), in which case it is easy to see that rotations by Euler angles: $(\alpha,\beta,\gamma) =(\phi_i,\theta_i,0)$ will align the cartesian system at $\hat{n}_0$ with that defined at position $\hat{n}_i$. 
%--------------------------------------------------------
\subsection{A heuristic argument for constructing spin-0 $E$ and $B$ fields}\label{sec:qu2eb_heuristic}

@@

 where the terms on either side of the equation are those that appear in \eq{eq:qu2eb_ker_1}. Note that each sub-matrix in \eq{eq:qu2eb_ker_1} is formed by some linear combination of a complex number and its conjugate and is hence real. The $m$ sum over the product of two spherical harmonic functions with spins $s_1$ and $s_2$ respectively, is given by the following identity \cite{varshalovich},

@@

Specifically note that $\alpha_{qe}, ~\beta_{qe} ~\&~ \gamma_{qe}$ denote the Euler angles which rotate the local cartesian system at the location $\hat{n}_q$ where the Stokes field is measured to the cartesian system at the location $\hat{n}_e$ where the scalar fields are evaluated: $\hat{n}_q \xrightarrow{\mathcal{R}(\alpha_{qe},\beta_{qe},\gamma_{qe})} \hat{n}_e$. The Euler angles for the inverse rotations are given by the Euler angles $\alpha_{eq}=-\gamma_{qe}$, $\beta_{eq} = -\beta_{qe}$ and  $\gamma_{eq} =-\alpha_{qe}$\cite{varshalovich}. Since the kernel only depends on the cosine of the Euler angle $\beta$, it is immune to changes in its sign. More importantly, the ordering of indices  on the Euler angle $\alpha$ and $\gamma$  is important and leads to two different interpretations of the real space kernels.

@@

\bar{S}_q = \bmat E_e \\ B_e  \emat_{q} =- {{}_{\mm}f}(\beta_{qe},\ell_{\rm min},\ell_{\rm max})\bmat \cos(2 \alpha_{q e}) & \sin(2\alpha_{q e})\\  -\sin(2 \alpha_{q e})  & \cos(2 \alpha_{q e}) \emat  \bmat Q_{q} \\ U_{q}  \emat \Delta \Omega \,.
